---
title: "Course Project Description"
name: "Jieying Li, 919349988  "
date: "3/16"
output: html_document
---
Name: Jieying Li, 919349988


# Abstract:

This project analyzes a subset of data that is collected from experiments conudcted by Steinmetz et al (2019), which highlights the 18 sessions involving four mice: Cori, Frossman, Hence, and Lederberg.This study had 10 mice per sessions and each session has multiple trails to show the stimuli of the mice. The mice decisions are based on these stimuli and their neural activity in the visual cortex was recorded. The purpose is to create a predictive model for trial outcomes using neural activity and stimulus contrast. The project is divided into 3 main parts: exploratory data analysis, data intergration, and model training and prediction. 
 

# Introduction:

Many neuoscience research have shown significant progress such as studies like the Steinnetz et al (2019) which provided datasets for analysis. This project focuses on 18 sub datasets, consists for recordings from the visual cortex of the mice duirng trials that involve visual stimuli. Each session comprises multiple trials where mice are presented with stimuli of varying contrast on two screen. Their neural activity and decisions are recorded to offer detials for the relationship between neural activity patterns, stimulus contrasts, and trial outcome. The project aim to develop predictive models for trial outcome using neural activity and stimulus contrast data. There are three parts of the project whcih include: exploratory data analysis, data integration, and model training and prediction. 

```{r echo=FALSE, eval=TRUE}

session=list()
for(i in 1:18){
  session[[i]]=readRDS(paste('./Data/session',i,'.rds',sep=''))
  print(session[[i]]$mouse_name)
  print(session[[i]]$date_exp)
  
}
```

#Exploratory Analysis:

```{r echo=TRUE, eval=TRUE, message=FALSE}

suppressWarnings(library(tidyverse))
suppressWarnings(library(knitr))
suppressWarnings(library(dplyr))
suppressWarnings(library(caret))
suppressWarnings(library(reshape2))
suppressWarnings(library(pROC))
suppressWarnings(library(xgboost))
```

```{r echo=FALSE, eval=TRUE}
library(tidyverse)

n.session=length(session)

meta <- tibble(
  mouse_name = rep('name',n.session),
  date_exp =rep('dt',n.session),
  n_brain_area = rep(0,n.session),
  n_neurons = rep(0,n.session),
  n_trials = rep(0,n.session),
  success_rate = rep(0,n.session)
)


for(i in 1:n.session){
  tmp = session[[i]];
  meta[i,1]=tmp$mouse_name;
  meta[i,2]=tmp$date_exp;
  meta[i,3]=length(unique(tmp$brain_area));
  meta[i,4]=dim(tmp$spks[[1]])[1];
  meta[i,5]=length(tmp$feedback_type);
  meta[i,6]=mean(tmp$feedback_type+1)/2;
  }

```

```{r echo=FALSE, eval=TRUE}
# In package knitr

kable(meta, format = "html", table.attr = "class='table table-striped'",digits=2) 

#Modify column names
new_column_names <- c("Mouse Name", "Date", "brain area","neurons","# trials", "% sucess rate")
names(meta) <- new_column_names 

# Generate HTML table using kable with modified column names
kable(meta, format = "html", table.attr = "class='table table-striped'", digits = 2)

```

(**)
```{r echo = FALSE}
dim(session[[1]]$spks[[1]]) 
length(session[[1]]$brain_area)
session[[1]]$spks[[1]][6,] # Each row contains 40 time bins. 

# connecting neuron spike with section 1 
session[[1]]$spks[[1]][6,3] 
session[[1]]$brain_area[6]
```

```{r echo = FALSE}
get_trail_data <- function(session_id, trail_id){
  spikes <- session[[session_id]]$spks[[trail_id]]
  if (any(is.na(spikes))){
    disp("value missing")
  }

  #trail_tibble <- as_tibble(spikes) %>% set_names(binename) %>%  add_column("brain_area" = session[[session_id]]$brain_area ) %>% group_by(brain_area) %>% summarize( "sum_spikes" =across(everything(),sum),.groups = "drop") 
  trail_tibble <- tibble("neuron_spike" = rowSums(spikes))  %>%  add_column("brain_area" = session[[session_id]]$brain_area ) %>% group_by(brain_area) %>% summarize( region_sum_spike = sum(neuron_spike), region_count = n(),region_mean_spike = mean(neuron_spike)) 
  trail_tibble  = trail_tibble%>% add_column("trail_id" = trail_id) %>% add_column("contrast_left"= session[[session_id]]$contrast_left[trail_id]) %>% add_column("contrast_right"= session[[session_id]]$contrast_right[trail_id]) %>% add_column("feedback_type"= session[[session_id]]$feedback_type[trail_id])
  trail_tibble
}

```

```{r echo =FALSE}
trail_tibble_1_2 <- get_trail_data(1,2)
trail_tibble_1_2
```

```{r echo = FALSE}
get_session_data <- function(session_id){
  n_trail <- length(session[[session_id]]$spks)
  trail_list <- list()
  for (trail_id in 1:n_trail){
    trail_tibble <- get_trail_data(session_id,trail_id)
    trail_list[[trail_id]] <- trail_tibble
  }
  session_tibble <- do.call(rbind, trail_list)
  session_tibble <- session_tibble %>% add_column("mouse_name" = session[[session_id]]$mouse_name) %>% add_column("date_exp" = session[[session_id]]$date_exp) %>% add_column("session_id" = session_id) 
  session_tibble
}

```

```{r echo = FALSE}
session_1 <- get_session_data(1)
head(session_1)
```

```{r echo = FALSE}
new_column_names <- c("Mouse Name", "Date", "brain area", "neurons", "# trials", "% success rate")
names(meta) <- new_column_names

# Extract relevant variables from meta dataframe
relevant_vars <- meta[, c("# trials", "% success rate")]


# Compute correlation matrix
cor_matrix <- cor(relevant_vars)

# Visualize correlation matrix using a heatmap
library(ggplot2)

# Convert correlation matrix to long format
cor_df <- reshape2::melt(cor_matrix)

# Plot heatmap
ggplot(cor_df, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0,
                       limit = c(-1,1), space = "Lab", name="Correlation") +
  theme_minimal() +
  labs(title = "Correlation Between Number of Trials and Success Rate")
```


From this graph, we can tell that the correlation before the success rate is moderately correlated, this means that as the number of trials increase, the success rate will also increase. 

```{r echo=FALSE}
i.s= 2 

i.t= 1 

spk.trial = session[[i.s]]$spks[[i.t]]
area=session[[i.s]]$brain_area

spk.count=apply(spk.trial,1,sum)

# for(i in 1:dim(spk.trial)[1]){
#  spk.count[i]=sum(spk.trial[i,])
# }


# tapply():
spk.average.tapply=tapply(spk.count, area, mean)


# dplyr: 
# create a data frame
tmp <- data.frame(
  area = area,
  spikes = spk.count
)
# Calculate the average by group using dplyr
spk.average.dplyr =tmp %>%
  group_by(area) %>%
  summarize(mean= mean(spikes))
```

```{r}
# Calculate success rate for each trial
success_rate_per_trial <- lapply(session, function(sess) {
  # Extract relevant variables for each session
  trials <- data.frame(
    contrast_left = unlist(sess$contrast_left),
    contrast_right = unlist(sess$contrast_right),
    feedback_type = unlist(sess$feedback_type)
  )
  
  # Define a function to determine success based on feedback type and contrasts
  determine_success <- function(left_contrast, right_contrast, feedback_type) {
    if (left_contrast > right_contrast && feedback_type == 1) {
      return(1)  # Success
    } else if (left_contrast < right_contrast && feedback_type == -1) {
      return(1)  # Success
    } else if (left_contrast == 0 && right_contrast == 0 && feedback_type == 1) {
      return(1)  # Success
    } else if (left_contrast == right_contrast && left_contrast != 0 && feedback_type == 1) {
      return(1)  # Success
    } else {
      return(0)  # Failure
    }
  }
  
  # Apply the determine_success function to each trial
  trials$success <- mapply(determine_success, trials$contrast_left, trials$contrast_right, trials$feedback_type)
  
  # Calculate success rate for each trial
  success_rate <- cumsum(trials$success) / seq_along(trials$success)
  
  # Return success rate for each trial
  return(success_rate)
})

# Plot success rate over time for individual sessions
plot_success_rate <- function(success_rate, session_number) {
  plot(success_rate, type = "l", xlab = "Trial Number", ylab = "Success Rate", 
       main = paste("Success Rate Over Time for Session", session_number))
}


# Plot success rate for each session
for (i in 1:length(success_rate_per_trial)) {
 
  plot_success_rate(success_rate_per_trial[[i]], i)
  
}
```

These plot shows the success rate for each trial within each session. I noticed that after the 50th trials, the graph tend to converge to a constant which in most cases it tend to be around 0.6.  

```{r}
#full_functional_tibble %>% group_by(mouse_name) %>% summarize(success_rate = mean(success, na.rm = TRUE))
```


```{r}
mouse_success_rates <- data.frame(
  mouse_name = character(),
  success_rate = double()
)

# Calculate success rate for each mouse
for (i in 1:length(success_rate_per_trial)) {
  mouse_name <- session[[i]]$mouse_name[1]
  success_rate <- tail(success_rate_per_trial[[i]], 1)  # Take the final success rate
  mouse_success_rates <- rbind(mouse_success_rates, data.frame(mouse_name = mouse_name, success_rate = success_rate))
}

# Plot success rate for individual mice
ggplot(mouse_success_rates, aes(x = mouse_name, y = success_rate, fill = mouse_name)) +
  geom_bar(stat = "identity") +
  labs(x = "Mouse Name", y = "Success Rate", title = "Success Rate for Individual Mice") +
  theme_minimal()
```

This graph shows the success rate of for each trial within each session and then aggregate the success rate by the mouse. From the graph, Leaderberg has the highest success rate. From table in the comment, Cori	0.6391231, Forssmann	0.6870813, Hench	0.6839121, Lederberg	0.7608268. The axis variable of the successful is not in percentage because 0-4, where 4 is the highest rank has a better presentation for the graph. 

```{r echo=FALSE} 

average_spike_area<-function(i.t,this_session){
  spk.trial = this_session$spks[[i.t]]
  area= this_session$brain_area
  spk.count=apply(spk.trial,1,sum)
  spk.average.tapply=tapply(spk.count, area, mean)
  return(spk.average.tapply)
  }

# Testing
average_spike_area(1,this_session = session[[i.s]])

```


This calculate teh average spike count for each brain area in a session. 

```{r echo=FALSE}

n.trial=length(session[[i.s]]$feedback_type)
n.area=length(unique(session[[i.s]]$brain_area ))


# create a data frame that contain the average spike counts for each area, feedback type,  the two contrasts, and the trial id

trial.summary =matrix(nrow=n.trial,ncol= n.area+1+2+1)
for(i.t in 1:n.trial){
  trial.summary[i.t,]=c(average_spike_area(i.t,this_session = session[[i.s]]),
                          session[[i.s]]$feedback_type[i.t],
                        session[[i.s]]$contrast_left[i.t],
                        session[[i.s]]$contrast_right[i.s],
                        i.t)
}

colnames(trial.summary)=c(names(average_spike_area(i.t,this_session = session[[i.s]])), 'feedback', 'left contr.','right contr.','id' )

# Turning it into a data frame
trial.summary <- as_tibble(trial.summary)
 
```



```{r echo=FALSE}
area.col=rainbow(n=n.area,alpha=0.7)

plot(x=1,y=0, col='white',xlim=c(0,n.trial),ylim=c(0.5,2.2), xlab="Trials",ylab="Average spike counts", main=paste("Spikes per area in Session", i.s))


for(i in 1:n.area){
  lines(y=trial.summary[[i]],x=trial.summary$id,col=area.col[i],lty=2,lwd=1)
  lines(smooth.spline(trial.summary$id, trial.summary[[i]]),col=area.col[i],lwd=3)
  }
legend("topright", 
  legend = colnames(trial.summary)[1:n.area], 
  col = area.col, 
  lty = 1, 
  cex = 0.8
)

```

<span style="color: blue;"> This plot shows the average spike counts for each neruon area across trials. This helps understand the activity level of different brain areas during the session 2. I noticed that the average spike count generally did not change as much as the numer of trails increase except for CA1 line and the root line. </span>

```{r echo=FALSE}

plot.trial<-function(i.t,area, area.col,this_session){
    
    spks=this_session$spks[[i.t]];
    n.neuron=dim(spks)[1]
    time.points=this_session$time[[i.t]]
    
    plot(0,0,xlim=c(min(time.points),max(time.points)),ylim=c(0,n.neuron+1),col='white', xlab='Time (s)',yaxt='n', ylab='Neuron', main=paste('Trial ',i.t, 'feedback', this_session$feedback_type[i.t] ),cex.lab=1.5)
    for(i in 1:n.neuron){
        i.a=which(area== this_session$brain_area[i]);
        col.this=area.col[i.a]
        
        ids.spike=which(spks[i,]>0) # find out when there are spikes 
        if( length(ids.spike)>0 ){
            points(x=time.points[ids.spike],y=rep(i, length(ids.spike) ),pch='.',cex=2, col=col.this)
        }
      
            
    }
    
legend("topright", 
  legend = area, 
  col = area.col, 
  pch = 16, 
  cex = 0.8
  )
  }
    
```

```{r, fig.width=8, fig.height=8}
varname=names(trial.summary);
area=varname[1:(length(varname)-4)]
plot.trial(1,area, area.col,session[[i.s]])
```


This graph shows the trial 1 feedback 1, showing the spike activity of the neurons. 
```{r, fig.width=8, fig.height=8}
varname=names(trial.summary);
area=varname[1:(length(varname)-4)]
par(mfrow=c(1,2))
plot.trial(1,area, area.col,session[[i.s]])
plot.trial(2,area, area.col,session[[i.s]])

par(mfrow=c(1,1))
```

<span> The two plots between trial 1 and 2 in feedback 1 shows some difference in spike activity. For VISpm, there are clearly more neurons in trial 1 than trial 2. The time for the neurons are higher in trial two than trial one. The difference in the time and the amount of neurons can suggest that there are confounding variables such as synaptic transmission delay, feedback loops, and networking connectivity. The changes in in network dynamics over time or across experimental conditions could potential influence difference results over time. When comparing the trials for feedback 1, the graphs shows that as the number of trials increase, the time increase and the neurons fired have decreased in some amount. 

```{r echo=FALSE}

sessions_data <- lapply(session[2:17], function(s) {
  data.frame(
    Feedback_Type = s$feedback_type,
    Contrast_Left = s$contrast_left,
    Contrast_Right = s$contrast_right
    
  )
})

# Combine data 
all_data <- do.call(rbind, sessions_data)

# Compute correlation matrix
cor_matrix <- cor(all_data)

# Visualize correlation matrix 
library(ggplot2)
library(reshape2)

cor_df <- melt(cor_matrix)
ggplot(cor_df, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0,
                       limit = c(-1,1), space = "Lab", name="Correlation") +
  theme_minimal() +
  labs(title = "Correlation Between Sessions 2 to 17")

```

This correlation matrix shows the feedback type, contrast left and right in the graph. We can see that there is weak correlation between contrast left and contrast right for the sessions 2 to 17. There are almost no correlation between feedback type and ocntrast right. 



# Data Integration 


```{r,echo=FALSE}
binename <- paste0("bin", as.character(1:40))

get_trail_functional_data <- function(session_id, trail_id){
  spikes <- session[[session_id]]$spks[[trail_id]]
  if (any(is.na(spikes))){
    disp("value missing")
  }

  trail_bin_average <- matrix(colMeans(spikes), nrow = 1)
  colnames(trail_bin_average) <- binename
  trail_tibble  = as_tibble(trail_bin_average)%>% add_column("trail_id" = trail_id) %>% add_column("contrast_left"= session[[session_id]]$contrast_left[trail_id]) %>% add_column("contrast_right"= session[[session_id]]$contrast_right[trail_id]) %>% add_column("feedback_type"= session[[session_id]]$feedback_type[trail_id])
  
  trail_tibble
}
get_session_functional_data <- function(session_id){
  n_trail <- length(session[[session_id]]$spks)
  trail_list <- list()
  for (trail_id in 1:n_trail){
    trail_tibble <- get_trail_functional_data(session_id,trail_id)
    trail_list[[trail_id]] <- trail_tibble
  }
  session_tibble <- as_tibble(do.call(rbind, trail_list))
  session_tibble <- session_tibble %>% add_column("mouse_name" = session[[session_id]]$mouse_name) %>% add_column("date_exp" = session[[session_id]]$date_exp) %>% add_column("session_id" = session_id) 
  session_tibble
}

```


```{r, echo = FALSE}
session_list = list()
for (session_id in 1: 18){
  session_list[[session_id]] <- get_session_functional_data(session_id)
}
full_functional_tibble <- as_tibble(do.call(rbind, session_list))
full_functional_tibble$session_id <- as.factor(full_functional_tibble$session_id )
full_functional_tibble$contrast_diff <- abs(full_functional_tibble$contrast_left-full_functional_tibble$contrast_right)

full_functional_tibble$success <- full_functional_tibble$feedback_type == 1
full_functional_tibble$success <- as.numeric(full_functional_tibble$success)
```


```{r}
predictive_feature <- c("session_id","trail_id","contrast_right","contrast_left", "contrast_diff" ,binename)
head(full_functional_tibble[predictive_feature])
```
This table shows the trails from all session and the variables name on the top is the average spike rate of each time bin. 


```{r}
predictive_dat <- full_functional_tibble[predictive_feature]
#predictive_dat$success <- as.numeric(predictive_dat$success)
predictive_dat$trail_id <- as.numeric(predictive_dat$trail_id)
label <- as.numeric(full_functional_tibble$success)
X <- model.matrix(~., predictive_dat)
```


```{r echo=FALSE}
#  calculate average spike rates across trials
calculate_average_spike_rates <- function(this_session) {
  n_trials <- length(this_session$spks)
  spike_rates <- numeric(n_trials)
  
  for (i.t in 1:n_trials) {
    spks <- this_session$spks[[i.t]]
    spike_counts <- apply(spks, 1, sum)
    spike_rates[i.t] <- mean(spike_counts)
  }
  
  return(spike_rates)
}

# Plot average spike rates across trials
plot_average_spike_rates <- function(this_session) {
  spike_rates <- calculate_average_spike_rates(this_session)
  
  plot(1:length(spike_rates), spike_rates, type = "l", 
       xlab = "Trial", ylab = "Average Spike Rate", 
       main = "Average Spike Rate Across Trials", col = "blue")
}

# Call the function to plot average spike rates
plot_average_spike_rates(session[[i.s]])

```


By examining the line plot, the overall trend in the spike rate are shown in above. The line plot shows that the average spike rate varies across the trials, therefore, we can not say there are noticeable trends. 




```{r echo=FALSE}

n.session=4

meta <- tibble(
  mouse_name = rep('name',n.session),
  date_exp =rep('dt',n.session),
  n_brain_area = rep(0,n.session),
  n_neurons = rep(0,n.session),
  n_trials = rep(0,n.session),
  success_rate = rep(0,n.session)
)

for(i in 1:4){
  tmp = session[[i]];
  meta[i,1]=tmp$mouse_name;
  meta[i,2]=tmp$date_exp;
  meta[i,3]=length(unique(tmp$brain_area));
  meta[i,4]=dim(tmp$spks[[1]])[1];
  meta[i,5]=length(tmp$feedback_type);
  meta[i,6]=mean(tmp$feedback_type+1)/2;
}
kable(meta, format = "html", table.attr = "class='table table-striped'",digits=2) 
```

```{r echo=FALSE}

spks.trial=session[[1]]$spks[[1]]
total.spikes=apply(spks.trial,1,sum)
(avg.spikes=mean(total.spikes))

# Average number of spikes per neuron in the first 0.4 seconds of Trial 1 in Session 1
spks_trial <- session[[1]]$spks[[1]]
total_spikes <- apply(spks_trial, 1, sum)
avg_spikes_first_trial <- mean(total_spikes)
print(avg_spikes_first_trial)

# Average number of spikes per active neuron
active_neurons <- sum(total_spikes > 0)
avg_spikes_per_active_neuron <- sum(total_spikes) / active_neurons
print(avg_spikes_per_active_neuron)

# Average number of spikes per neuron in Brain area X
brain_area_X <- "TH"  # Replace "X" with the desired brain area
neurons_in_area_X <- which(session[[1]]$brain_area == brain_area_X)
spikes_in_area_X <- total_spikes[neurons_in_area_X]
avg_spikes_per_neuron_in_area_X <- mean(spikes_in_area_X)
print(avg_spikes_per_neuron_in_area_X)

average_spike_table <- data.frame(
  Metric = c("First Trial in Session 1", "Per Active Neuron", paste("Per Neuron in", brain_area_X)),
  Average_Spike_Count = c(avg_spikes_first_trial, avg_spikes_per_active_neuron, avg_spikes_per_neuron_in_area_X)
)

# Print the table
kable(average_spike_table, format = "html", 
      table.attr = "class='table table-striped'",
      digits = 2, 
      col.names = c("Metric", "Average Spike Count"))

```
This table shows the avergae number of spikes per neuron in the first 0.4 seconds of trial 1 in session 1, average number of spikes per active neuron and the average number of spikes per neurons in the specified brain area.


# Model Training and Prediction 

```{r echo=FALSE}
n_obs = length(session[[18]]$feedback_type)

dat = tibble(
    feedback_type = as.factor(session[[18]]$feedback_type),
    decision = rep('name', n_obs),
    avg_spikes = rep(0, n_obs)
)

for (i in 1:n_obs){
    # decision 
    if (session[[18]]$contrast_left[i] > session[[18]]$contrast_right[i]){
        dat$decision[i] = '1' 
    } else if (session[[18]]$contrast_left[i] < session[[18]]$contrast_right[i]){
        dat$decision[i] = '2' 
    } else if (session[[18]]$contrast_left[i] == session[[18]]$contrast_right[i] 
               & session[[18]]$contrast_left[i] == 0){
        dat$decision[i] = '3' 
    } else{
        dat$decision[i] = '4' 
    }
    
    # avg_spks
    spks.trial = session[[18]]$spks[[i]]
    total.spikes = apply(spks.trial, 1, sum)
    dat$avg_spikes[i] = mean(total.spikes)
}

dat$decision = as.factor(dat$decision)

summary(dat)
```

This table tells us about the characteristics of the feedback type, decisions, and the average spikes varaibke in session 18. 

```{r}
# Split data into train and test
set.seed(101)
sample <- sample.int(n = n_obs, size = floor(.8 * n_obs), replace = F)
train <- dat[sample, ]
test  <- dat[-sample, ]
```

```{r echo=FALSE}
fit1 <- glm(feedback_type~., data = train, family="binomial")
summary(fit1)
```

```{r echo=FALSE}
pred1 <- predict(fit1, test %>% select(-feedback_type), type = 'response')
prediction1 <- factor(pred1 > 0.5, labels = c('-1', '1'))
mean(prediction1 != test$feedback_type)
```

The shows that the prediction error is about 23% 


```{r echo=FALSE}
average_spike_area<-function(i.t,this_session){
  spk.trial = this_session$spks[[i.t]]
  area= this_session$brain_area
  spk.count=apply(spk.trial,1,sum)
  spk.average.tapply=tapply(spk.count, area, mean)
  return(spk.average.tapply)
}
```

```{r echo=FALSE}
n_area = length(unique(session[[18]]$brain_area))
spk_area = matrix(rep(0, n_obs * n_area), n_obs, n_area)
for (i in 1:n_obs){
    spk_area[i,] = average_spike_area(i, session[[18]])
}

spk_area = as_tibble(spk_area)
colnames(spk_area)= unique(session[[18]]$brain_area)
dat1 = bind_cols(dat, spk_area) %>% select(-avg_spikes)
head(dat1)
```


This table tells us the average spike rate for each neuron in a given session 18. 
```{r echo=FALSE}
# Split data into train and test
set.seed(101)
sample <- sample.int(n = n_obs, size = floor(.8 * n_obs), replace = F)
train <- dat1[sample, ]
test  <- dat1[-sample, ]
```

```{r echo=FALSE}
fit2 <- glm(feedback_type~., data = train, family="binomial")
summary(fit2)
```

```{r echo=FALSE}
pred2 <- predict(fit2, test %>% select(-feedback_type), type = 'response')
prediction2 <- factor(pred2 > 0.5, labels = c('-1', '1'))
mean(prediction2 != test$feedback_type)
```
This prediction model shows a prediction error of 25% which is worst than the result we have gotten earlier. This means (insert)




```{r MODELING}
# split
set.seed(123) # for reproducibility
trainIndex <- createDataPartition(label, p = .8, 
                                  list = FALSE, 
                                  times = 1)
train_df <- predictive_dat[trainIndex, ]
train_X <- X[trainIndex,]
test_df <- predictive_dat[-trainIndex, ]
test_X <- X[-trainIndex,]

train_label <- label[trainIndex]
test_label <- label[-trainIndex]
```

```{r}
library(xgboost)
xgb_model <- xgboost(data = train_X, label = train_label, objective = "binary:logistic", nrounds=10)

```

I decieded to use xgboost because I have a lot of features, similarly to the Ta's demo. The code demonstrates the process of splitting the data into training and test sets for binary classification.

```{r}
predictions <- predict(xgb_model, newdata = test_X)
predicted_labels <- as.numeric(ifelse(predictions > 0.5, 1, 0))
accuracy <- mean(predicted_labels == test_label)
accuracy

```
```{r}
conf_matrix <- confusionMatrix(as.factor(predicted_labels), as.factor(test_label))
conf_matrix$table

```
```{r}
#install.packages("pROC")  # Install the pROC package
library(pROC)  
auroc <- roc(test_label, predictions)
auroc
```

The above shows the prediction result for accuracy, confusion matrix, and auroc. Auroc helpes measure the performance of a binary classification model. Area under the curve is 0.7379 which means that our model with moderate discrimination ability and it suggest that there are some predictive power. 


#This tests the data from session 18 on performances of 50 random trails. 
```{r}

# split
set.seed(123) # for reproducibility
session_18_row <- which(full_functional_tibble$session_id==18)
testIndex <- sample(session_18_row, 50, replace = FALSE)
trainIndex <- 1:nrow(full_functional_tibble)
trainIndex <- trainIndex[!(trainIndex %in% testIndex)]

train_df <- predictive_dat[trainIndex, ]
train_X <- X[trainIndex,]
test_df <- predictive_dat[-trainIndex, ]
test_X <- X[-trainIndex,]

train_label <- label[trainIndex]
test_label <- label[-trainIndex]
```


```{r}
xgb_model <- xgboost(data = train_X, label = train_label, objective = "binary:logistic", nrounds=10)
predictions <- predict(xgb_model, newdata = test_X)
predicted_labels <- as.numeric(ifelse(predictions > 0.5, 1, 0))
accuracy <- mean(predicted_labels == test_label)
accuracy
conf_matrix <- confusionMatrix(as.factor(predicted_labels), as.factor(test_label))
conf_matrix$table
auroc <- roc(test_label, predictions)
auroc
```

With an area under the curve of 0.6875, it suggest that our model has some discriminatory power and there can be room for improvement. 


## This test the model's performance on 50 random trails from session 1
```{r,echo=FALSE}
# split
set.seed(123) # for reproducibility
session_1_row <- which(full_functional_tibble$session_id==1)
testIndex <- sample(session_1_row, 50, replace = FALSE)
trainIndex <- 1:nrow(full_functional_tibble)
trainIndex <- trainIndex[!(trainIndex %in% testIndex)]

train_df <- predictive_dat[trainIndex, ]
train_X <- X[trainIndex,]
test_df <- predictive_dat[-trainIndex, ]
test_X <- X[-trainIndex,]

train_label <- label[trainIndex]
test_label <- label[-trainIndex]
```

```{r,echo=FALSE}
xgb_model <- xgboost(data = train_X, label = train_label, objective = "binary:logistic", nrounds=10)
predictions <- predict(xgb_model, newdata = test_X)
predicted_labels <- as.numeric(ifelse(predictions > 0.5, 1, 0))
accuracy <- mean(predicted_labels == test_label)
accuracy
conf_matrix <- confusionMatrix(as.factor(predicted_labels), as.factor(test_label))
conf_matrix$table
auroc <- roc(test_label, predictions)
auroc
```


With an area under the curve of 0.6689, it suggest that our model has some discriminatory power and there can be room for improvement. 

#Conclusion

In the exploratory data analysis, I looked at neural and behavior data obtained from series of experiments with the 4 different mice. By looking at the characteristics of the dataset across multiple sessions, it gave us more information and the trends for the data. The changes across trials within sessions shows the success rates varied over time. 

In the data integration, I made an approach to combine data from multiple trials so that I can extract shared patterns and address differences between the sessions. 

For the last part, I trained a prediction model to predict the outcome base on neural activity and stimuli information. This was an interesting project to look at. 

# Reference {-}


(**) Side note, there are two of the same graph because I have changed the name of the 1st graph to the second. 


Steinmetz, N.A., Zatka-Haas, P., Carandini, M. et al. Distributed coding of choice, action and engagement across the mouse brain. Nature 576, 266–273 (2019). https://doi.org/10.1038/s41586-019-1787-x

TA and professor's code. 


